### 2024-04-04 manuscript.Rmd
update

Compared to other snakes, the vertebral column of fossorial snakes tend to feature a much higher proportion of pre- than post-cloacal segments [@pollyTesting2001]. This is reflected in the relative short tails of blindsnakes (CITE).

@@

3. Do vertebral number and body size vary with mean annual temperature?

@@

<!-- ## Environmental data  -->
<!-- NEED TO PUT IN THE CORRECT VALUES HERE -->
<!-- We used the Australian blindsnake occurrence records presented in (Tiatragul et al. unpublished). These records were compiled from museum databases and verified human observations from Atlas of the Living Australia (ALA). A total of XXXX records (min = , median = , max = , median per species) representing XX species of Australian *Anilios* for our analyses.

@@

All statistical analyses were performed in R v.4.31 [@rcoreteamLanguage2023]. To test for evolutionary correlations between the number of vertebrae and body size and proportions, we used phylogenetic generalised least squares (PGLS) regression under a model of Browninan motion implemented in `geomorph` v.4.0.6 [@adamsGeomorph2023]. First, we tested for pleomerism [@lindseyPleomerism1975] across 37 species of *Anilios* blindsnakes. 
We did this by testing whether the total number of vertebrae positively correlate with body size when regressing the maximum number of total vertebrae against maximum total body length from Tiatragul et al. (unpublished). Then, we test whether the vertebrae to body length ratio is related to maximum aspect ratio, defined by total body length / body width, also from Tiatragul et al. (unpublished) data.

@@

### Relationship between pleomerism and body shape
However, we found that vertebrae ratio is postively correlated with aspect ratio (PGLS, slope=-0.0048, y=0.73, R^2^=0.17, F=7.09,, Z=2.2, P=0.011), indicating that species that are more narrow-and-wide usually have greater total vertebrae to total length ratios than species that are wide-and-robust (Fig. \@ref(fig:fig-vert-width-ratios)). This is consistent with colubrids and viperids, where species that are more elongated tend to have greater number of vertebrae than stouter ones [@lindellEvolution1994].

@@

<!-- ### Variation in vertebrae number -->

Closely related species tend to have similar number of total vertebrae. 
Males tend to have greater number of postcloacal vertebrae

<!-- FIGURE 1 -->

There is a statistically significant correlation between vertebrae ratio and fineness of the body. Species that have more vertebrae per length (vertebrae ratio > 1) are usually more fine, whereas more robust species have lower number of vertebrae.

@@

Temperature seem to have XXX effect. Common garden experiments for different species of snakes can tell us if incubation temperature and thus rate of development influences number of vertebrae. We would expect higher temperature to accelerate development (CITE anolis papers, turtles, etc), so less time for embryonic development.

>>>>

Compared to other snakes, the vertebral column of fossorial snakes tend to feature a much higher proportion of pre- than post-cloacal segments [@pollyTesting2001]. This is reflected in the relatively short tails of blindsnakes (CITE).

@@

3. Does vertebral ratio vary with soil compactness or mean annual temperature?

@@

## Environmental data 
<!-- NEED TO PUT IN THE CORRECT VALUES HERE -->

We used the Australian blindsnake occurrence records presented in (Tiatragul et al. unpublished). These records were compiled from museum databases and verified human observations from Atlas of the Living Australia (ALA). A total of XXXX records (min = , median = , max = , median per species) representing XX species of Australian *Anilios* for our analyses.

<!-- We used the Australian blindsnake occurrence records presented in (Tiatragul et al. unpublished). These records were compiled from museum databases and verified human observations from Atlas of the Living Australia (ALA). A total of XXXX records (min = , median = , max = , median per species) representing XX species of Australian *Anilios* for our analyses.

@@

All statistical analyses were performed in R v.4.31 [@rcoreteamLanguage2023]. To test for evolutionary correlations between the number of vertebrae and body size and proportions, we used phylogenetic generalised least squares (PGLS) regression under a model of Browninan motion implemented in `geomorph` v.4.0.6 [@adamsGeomorph2023]. First, we tested for pleomerism [@lindseyPleomerism1975] across 37 species of *Anilios* blindsnakes. 
We did this by testing whether the total number of vertebrae positively correlate with body size when regressing the maximum number of total vertebrae against maximum total body length from Tiatragul et al. (unpublished). Next, we test whether the vertebrae to body length ratio is related to maximum aspect ratio, defined by total body length / body width, also from Tiatragul et al. (unpublished) data. Finally, we tested if variation in vertebral ratio can be explained by mean annual temperature.

@@

### Relationship between vertebrae ratio and body shape
However, we found that vertebrae ratio is postively correlated with aspect ratio (PGLS, slope=-0.0048, y=0.73, R^2^=0.17, F=7.09,, Z=2.2, P=0.011), indicating that species that are more narrow-and-wide usually have greater total vertebrae to total length ratios than species that are wide-and-robust (Fig. \@ref(fig:fig-vert-width-ratios)). This is consistent with colubrids and viperids, where species that are more elongated tend to have greater number of vertebrae than stouter ones [@lindellEvolution1994].

@@

We found a positive relationship between vertebrae ratio and species mean annual temperature (PGLS, slope=0.038, y=0.19, R^2^=0.21, F=9.58,, Z=2.32, P=0.009), indicating species that have greater amount of vertebrae for body size are generally found in places that are warmer.

knitr::include_graphics(paste0(WD, '/assets/figures/vert-ratio-to-temperature.jpg'))
```

@@

Mean annual temperature was a significant predictor of vertebrae ratio. Temperature could be a proxy for several ecological factors such ...
Common garden experiments for different species of snakes can tell us if incubation temperature and thus rate of development influences number of vertebrae. We would expect higher temperature to accelerate development (CITE anolis papers, turtles, etc), so less time for embryonic development.
